---
title: "Take-home Ex3: Visual Statistical Analysis of Resale Flats in 2022"
date: "February 10, 2023"
date-modified: last-modified
execute:
  echo: true
  eval: true
  warning: false
---

Installing and launching R packages

```{r}
pacman::p_load(ggstatsplot, plotly, performance, corrplot, tidyverse)
```

Loading the data

```{r}
resale <- read.csv("data/resale.csv")
```

Filtering the data to the required scope of 3/4/5-ROOM types and for the time period of 2022

```{r}
filter1 <- c("3 ROOM", "4 ROOM", "5 ROOM")
resale2022 <- filter(resale, resale$flat_type %in% filter1)

filter2 <- c("2022-01", "2022-02", "2022-03", "2022-04", "2022-05", "2022-06", "2022-07", "2022-08", "2022-09", "2022-10", "2022-11", "2022-12")
resale2022 <- filter(resale2022, resale2022$month %in% filter2)
```

Calculating the price per sqm

```{r}
resale2022 <- resale2022 %>% mutate(price_sqm = resale_price / floor_area_sqm)
```

Check distribution of data

```{r}
ggplot(data = resale2022,
       aes(x = price_sqm)) +
  geom_histogram(bins = 50) +
  facet_wrap(~ flat_type)
```

Using inverse to transform data with severe positive skew to fit normal distribution

```{r}
resale2022 <- resale2022 %>% mutate(adj_ppsm = 1/price_sqm)

ggplot(data = resale2022,
       aes(x = adj_ppsm)) +
  geom_histogram(bins = 50) +
  facet_wrap(~ flat_type)
```

One-way ANOVA test on Resale Prices by flat type

```{r}
ggbetweenstats(
  data = resale2022,
  x = flat_type,
  y = adj_ppsm,
  type = "p",
  mean.ci = TRUE,
  pairwise.comparisons = TRUE,
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Scatterplot

```{r}
plot_ly(data = resale2022,
        x = ~price_sqm,
        y = ~floor_area_sqm,
        color = ~flat_type)
```

Calculate lease term left

```{r}
resale2022 <- resale2022 %>% 
  mutate(tenure = substr(remaining_lease, start = 1, stop = 2)) %>% 
  mutate(tenure = as.numeric(tenure))
```

```{r}
plot_ly(data = resale2022,
        x = ~price_sqm,
        y = ~tenure)
```

Price predictions

```{r}
model <- lm(resale_price ~ floor_area_sqm, lease_commence_date, tenure, 
            data = resale2022)
model
```

```{r}
#check_c <- check_collinearity(model)
#plot(check_c)
```

```{r}
check_n <- check_normality(model)
plot(check_n)
```
