---
title: "In-class Ex5c: Heat Maps"
date: "February 8, 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
---

## **Heat Maps**

Installing and launching R packages, and loading data

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
wh <- read_csv("data/WHData-2018.csv") 
```

### **Static Heatmaps**

Preparing the data using by changing the rows by country name instead of row number

```{r}
row.names(wh) <- wh$Country
```

Transforming the data frame into a matrix to be able to create heat maps

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

Plot default cluster heatmap using heatmap() of base R Stats package

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

Use the arguments Rowv=NA and Colv=NA to switch off the option of plotting the row and column dendrograms.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

To normalize the matrix using scale argument for a more informative visual. Also note that margins argument is used to ensure that the entire x-axis labels are displayed completely and, cexRow and cexCol arguments are used to define the font size used for y-axis and x-axis labels respectively.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

### **Interactive Heatmaps**

Using heatmaply package

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

Scaling method when assume to be normal distribution

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

Normalization method when assume to be different or non-normal distributions so that it is easily comparable on the same scale

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

Percentizing method

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

OLO - optimal leaf ordering, GW - Gruvaeus and Wainer

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

Other plotting features to ensure cartographic quality heatmap can be produced

-   *colors* is used to change the colour palette

-   *k_row* is used to produce 5 groups.

-   *margins* is used to change the top margin to 60 and row margin to 200.

-   *fontsizw_row* and *fontsize_col* are used to change the font size for row and column labels

-   *main* is used to write the main title of the plot.

-   *xlab* and *ylab* are used to write the x-axis and y-axis labels respectively.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
